%\TODO{Introduction mentions interoperability between two OpenMP runtimes. But this section does not mention this. That is part of use case #2}

\subsection{Use Cases of OpenMP Interoperability}
We have identified four use cases of OpenMP interoperating with itself and other parallel programming models. 
\subsubsection{Interacting With User threads}
\input{userthread}

\subsubsection{Interacting with Other OpenMP Programs}
\input{ompomp}

\subsubsection{Interacting With Other Parallel Systems or Libraries}
\input{otherruntime}

\subsubsection{Interoperability with inter-node programming model, e.g. MPI}
\input{internode}

%\subsubsection{OpenMP with native threads}
%\subsubsection{OpenMP with TBB/MKL}
%\subsubsection{OpenMP with parallel scientific library, such as MKL}
%\subsubsection{Linking libraries and objects built with different OpenMP compilers}
%\subsubsection{OpenMP with inter-node model, e.g. MPI}


\subsection{Issues with Poor Interoperability}
\subsubsection{CPU Oversubscription}
\input{oversubscription}

\subsubsection{Conflicting Thread Affinity}
\input{threadaffinity}

\subsubsection{Global Impact of Environment Variables}
\input{environmentvariable}


\subsection{Limitation of Interoperability Support in the Standard}
OpenMP currently (4.0) provides limited support for users to give hints to runtime for 
better managing OpenMP threads and native threads, which can be used to help reducing 
the impact of oversubscriptions.
\paragraph{OMP\_DYNAMIC} % environment variable}
This also includes dynvar ICV, omp\_set\_dynamic and omp\_get\_dynamic runtime routine. OMP\_DYNAMIC could be either
{\bf true} or {\bf false}. When setting the dynvar ICV to be {\bf true}, user will expect
OpenMP implementation to adjust the number of threads to use for executing parallel
regions in order to optimize the use of system resources.

\paragraph{OMP\_WAIT\_POLICY} %environment variable}
This also include waitpolicyvar ICV, but no getter and setter routine. 
OMP\_WAIT\_POLICY could be set as either ACTIVE or PASSIVE. 
The ACTIVE value specifies that waiting threads should mostly be active, consuming processor
 cycles, while waiting. An OpenMP implementation may, for example, make waiting threads spin.
 The PASSIVE value specifies that waiting threads should mostly be passive, not consuming
 processor cycles, while waiting. For example, an OpenMP implementation may make waiting
 threads yield the processor to other threads or go to sleep.
 The details of the ACTIVE and PASSIVE behaviors are implementation defined.
 
 \paragraph{OMP\_THREAD\_LIMIT}
 This also include threadlimitvar ICV and omp\_get\_thread\_limit getter runtime routine.
 The environment variable sets the maximum number of OpenMP threads to use in a contention group by setting the threadlimitvar ICV.
The behavior of the program is implementation defined if the requested value of OMP\_THREAD\_LIMIT is greater than the number of threads an implementation can support

OMP\_DYNAMIC and OMP\_THREAD\_LIMIT are approaches to
addressing active oversubscriptions, and OMP\_WAIT\_POLICY could be used to address
passive oversubscriptions. 
Since there are no setters for ICVs for OMP\_WAIT\_POLICY
and OMP\_THREAD\_LIMIT variable in the current standard, 
dynamically changing waiting policy and the maximum number of
threads at runtime is not available.
