The complete OMP\_DYNAMIC spec reads: 
``The OMP\_DYNAMIC environment variable controls dynamic adjustment of the number of threads
to use for executing parallel regions by setting the initial value of the dyn-var ICV. The value of
this environment variable must be true or false. If the environment variable is set to true, the
OpenMP implementation may adjust the number of threads to use for executing parallel
regions in order to optimize the use of system resources. If the environment variable is set to
false, the dynamic adjustment of the number of threads is disabled. The behavior of the program
is implementation defined if the value of OMP\_DYNAMIC is neither true nor false.''


The complete OMP\_WAIT\_POLICY spec reads: 
``The OMP\_WAIT\_POLICY environment variable provides a hint to an OpenMP
 implementation about the desired behavior of waiting threads by setting the wait-policy-var ICV. A
 compliant OpenMP implementation may or may not abide by the setting of the environment
 variable.
 The value of this environment variable takes the form:
 ACTIVE \textbar  PASSIVE. 
 The ACTIVE value specifies that waiting threads should mostly be active, consuming processor
 cycles, while waiting. An OpenMP implementation may, for example, make waiting threads spin.
 The PASSIVE value specifies that waiting threads should mostly be passive, not consuming
 processor cycles, while waiting. For example, an OpenMP implementation may make waiting
 threads yield the processor to other threads or go to sleep.
 The details of the ACTIVE and PASSIVE behaviors are implementation defined.''

 
The complete OMP\_THREAD\_LIMIT spec reads:
``The OMP\_THREAD\_LIMIT environment variable sets the maximum number of OpenMP threads to use in a contention group by setting the thread-limit-var ICV.
The value of this environment variable must be a positive integer. The behavior of the program is implementation defined if the requested value of OMP\_THREAD\_LIMIT is greater than the number of threads an implementation can support, or if the value is not a positive integer.''
