\documentclass{article}
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf}

\headheight 0in
\oddsidemargin 0in
\evensidemargin  0in
\topmargin  -.25in
\textwidth 6.5in
\textheight 9in
%\title{OMPinterop: Interoperability and Composability of OpenMP\textsuperscript{\textregistered}}
\title{Interoperability and Composability of OpenMP\textsuperscript{\textregistered}}
\author{Yonghong Yan\thanks{cOMPunity}, 
%John Mellor-Crummey\thanks{Rice University}, 
%Martin Schulz\thanks{Lawrence Livermore National Laboratory},
\\~\\
%Nawal Copty\thanks{Oracle}, 
%Jim Cownie\thanks{Intel},
% John DelSignore\thanks{Rogue Wave}, 
%Robert Dietrich\thanks{TU Dresden, ZIH},
%Xu Liu\hbox to 0in{$^\dagger$\hss},
%Eugene Loh\hbox to 0in{$^\S$\hss}, 
%Daniel Lorenz\thanks{J\"{u}lich Supercomputer Center}, 
\\
and other members of the OpenMP Interoperability Subcommittee}

\usepackage{comment}
\usepackage{needspace}
\usepackage[colorlinks=true,citecolor=blue]{hyper ref}
\usepackage{url}
\usepackage{xcolor}

% correct bad hyphenation here
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{fancyvrb}
\usepackage{amsmath, amssymb}
\usepackage{epsfig}
\usepackage{setspace}\usepackage{multirow}
\usepackage{subfigure}
\usepackage{verbatim}
\usepackage{listings}
\usepackage{wrapfig}
%\usepackage{hyperref}
%\usepackage{xltxtra}
\newcommand{\REM}[1]{}

\newcommand{\descheader}[1]{{\needspace{3\baselineskip}\vspace{1em}\noindent \fbox{#1}}}


\begin{document}                                                
\maketitle
\section{Introduction}
Parallel and large-scale applications are typically developed using multiple parallel programming interfaces in 
a hybrid model, e.g. MPI+OpenMP, and using one or mulitple prebuilt scientific and/or platform-specific libraries such as MKL.
Each of these programming models and libraries has their own runtime to handle scheduling of work units and management of computational
and data movement tasks. There have been challening issues for using these models in one application, including 
compatibility issues for compiling and linking, oversubscription of resources at runtime, and the naming conflicts that 
programmers have to create workaround wrappers to deal with.

This report propose solutions to the interoperability and composability challenges faced by OpenMP programming interface, includling those
between multiple OpenMP implementations and/or multiple OpenMP runtime instances of the same implementation, OpenMP 
with native threads (pthreads and Windows Native threads), OpenMP with other threading languages and library such 
as C++11, TBB and Cilkplus, and OpenMP with inter-node programming models such as MPI, PGAS implemenation, etc. We 
think the similar challenges exist in other threading based libraries and language implementations, and believe
the solutions we provided in this technical report will work for them too.  

Interoperability and composability are closely related, while the interoperability sounds to improve the interactions between multiple models
while composability is meant to improve the modular use of OpenMP with itself and other models. One is from the aspect of system while 
the other is more concerned with software engineering. Both should be considered when developing solutions. 

For parallel programming languages and libraries, most implementations rely on system native threading (pthread or Windows Native threads) 
mechanisms to acquires system resources. Each implementation of the same or different programming models has their own mechanism for scheduling
user-level tasks and operations, which is the core part of a runtime system. 
The interoperability challenges are then concerned with how much we want two or more 
runtime instances (for the same or different high-level programming interfaces) to interact with other other for computational resource sharing
and data movement. Thus solutions to these challenges are more in the scope of runtime and implementation, than in the level of programming
interfaces and compiler transformations. 

\section{Topics and Interests}

\subsection{User threads}
\input{userthread}

\subsection{Oversubscription}
\input{oversubscription}

\subsection{Interaction between contention groups}

\subsection{Affinity with user threads}
\subsubsection{Coherence group (domains)}

\subsection{Common Interface}
\subsubsection{Application Binary Interfaces(ABI)}

\subsection{Interoperability with node-level programming model, e.g. MPI}

\begin{comment}
\subsection{Use Cases}
We have identified several use cases of OpenMP interoperating with itself and other parallel programming models. 
\subsubsection{OpenMP with native threads}
\subsubsection{OpenMP with TBB/MKL}
\subsubsection{OpenMP with parallel scientific library, such as MKL}
\subsubsection{Linking libraries and objects built with different OpenMP compilers}
\subsubsection{OpenMP with inter-node model, e.g. MPI}

\subsection{Issues and Posibble Solutions}
\subsubsection{Active Oversubscription}
\subsubsection{Passive Oversubscription}
\subsubsection{Interop across contention group}
\end{comment}

\section*{Acknowledgments}

The authors would like to acknowledge A, B, C and \ldots

\clearpage
\bibliographystyle{abbrv}
\bibliography{interop}

\appendix
\clearpage

\section{OpenMP 4.1 Formal specification}
\input{ompspec}

%\clearpage

\end{document}

